### 2016-04-26 p00-ch00-intro/intro.tex
polish

\paragraph*{Existing Concurrency Models}
In a concurrent setting, coroutines~\cite{conway1963design,taocp:knuth} enable interactions with collaborative pre-emption:
a coroutine has no \emph{return} statement, but it may explicitly \emph{yield}
to another coroutine (transfer control of the execution).
The yield relation is symmetric. 
Creating an instance of a coroutine spawns a new process with its own state.
A coroutine allows multiple entry points for suspending and resuming execution at explicitly specified locations.
The explicit interleaving points supports a compositional understanding of the code~\cite{DBLP:journals/cacm/Hoare74}.
coroutines are not originally established in the object-oriented paradigm.
Object orientation is based on caller-callee interaction, which is asymmetric:
the caller invokes a method in the callee and \emph{blocks} until the
corresponding \emph{return} occurs in the method of the callee.
Object orientation focuses on an object-view with caller-callee interaction and stack-based execution whereas coroutines focus on a process-view with flexible transfer of control.

@@

ABS~\cite{johnsen2012abs,hahnlehjlssw11} is a modeling language for concurrent objects and distributed systems.
ABS uses asynchronous communication among objects.
A message in ABS is bound to a method invocation;
this defines the interface of the sent messages.
In addition, ABS supports future values in its asynchronous communication.
ABS introduces \bfjtt{release} semantics that is based on co-operative scheduling of objects; 
i.e. similar to that of \emph{yield} in coroutines~\cite{creol:broch:owe}.
The ABS semantics is completely formalized by a structural operational semantics~\cite{plotkin:sos,abs:2012}.
This allows ABS models to take advantage of a wide range of static and dynamic analysis techniques~\cite{DBLP:journals/sttt/WongBBGGHMS15,DBLP:conf/cade/DinBH15,DBLP:conf/ifm/GiachinoGLLW13,DBLP:conf/tacas/AlbertAFGGMPR14}.
% 
Co-operative scheduling in ABS has been additionally extended for real-time scheduling with priorities and time constraints~\cite{bjork2013:rtabs,johnsen2012modeling}.
All the above characteristics make the ABS language an attractive choice if it
can be used as a programming language at industrial and business scale.
ABS has mainly developed as a modeling language;
for example, ABS does not support common I/O API for files or network,
and it does not provide a standard library of data structures.

JSR 166~\cite{jsr166} introduced a new concurrency model~\cite{DBLP:conf/java/Lea00} in Java 1.5.
The concurrency model lays the grounds to support actor model with asynchronous method calls.
This gave rise to various research to support actor model on Java~\cite{actor_frameworks_jvm:agha} and further extend Java as a functional programming language~\cite{odersky1997pizza,henkel2003discovering,nystrom2003polyglot,bracha1998making}.
From an industrial perspective, the development of Java language from version 6 to version 8 experienced a slow pace that influenced the growth of businesses and justified alternatives.
Scala~\cite{odersky2004scala}, a dynamic, functional, and object-oriented language on JVM, rose to the challenge to fill the gap during the period between Java~7 and Java~8.
Java~8~\cite{jls8} released fundamental features such as lambda expressions that are essential for concurrency and functional programming.

@@

\paragraph{Scalability} 
In distributed systems, the number of massages delivered among objects in the 
environment is not predictable at runtime.
The goal is to ensure the actor system scales in performance with least 
influence from the number of asynchronous messages delivered in the system.
Due to support of co-operative scheduling for asynchronous messages, ABS is a fit for distributed systems.

\paragraph{Separation of Concerns} 
We approach ABS modeling and development with a component-based or modular 
software engineering practices.
The scope of the research spans a number of layers around ABS language:

@@

The core contributions of this thesis target the intersection of object 
orientation, actor model, and concurrency.
We choose Java as the main programming language and as one of the mainstream 
object-oriented languages. 
We formalize a subset of Java and its concurrency API~\cite{jsr166} to 
facilitate formal verification and reasoning about it (cf. Chapter~\ref{ch:p01:ch02}).
We create an abstract mapping from a concurrent-object modeling language, 
ABS~\cite{johnsen2012abs}, to the programming semantics of concurrent Java (cf. Chapter~\ref{ch:p01:ch02}). 
We provide the formal semantics of the mapping and runtime properties of 
the concurrency layer including deadlines and scheduling policies (cf. Chapter~\ref{ch:p01:ch01}).
We provide an implementation of ABS concurrency layer as a Java API library 
and framework utilizing the latest language additions 
in Java 8~\cite{jsr335:lambda:translation} (cf. Chapter~\ref{ch:p02:ch01}).
We design and implement a runtime monitoring framework, JMSeq, to verify the
the correct ordering of execution of methods through code annotations in JVM (cf. Chapter~\ref{ch:p03:ch02}). 
In addition, we design a large-scale monitoring system as a real-world 
application; the monitoring system is built with ABS concurrent objects 
and formal semantics that leverages schedulability 
analysis to verify correctness of the monitors~\cite{fersman2007task} (cf. Chapter~\ref{ch:p03:ch01}).
Table~\ref{tbl:thesis} summarizes the structure of this text.

>>>>

\paragraph*{Existing Concurrency Models}
In a concurrent setting, coroutines~\cite{conway1963design,taocp:knuth} enable interactions with collaborative pre-emption:
a coroutine has no \emph{return} statement, but it may explicitly \emph{yield}
to another coroutine (transfer control of the execution).
The yield relation is symmetric. 
Creating an instance of a coroutine spawns a new process with its own state.
A coroutine allows multiple entry points for suspending and resuming execution at explicitly specified locations.
The explicit interleaving points supports a compositional understanding of the code~\cite{DBLP:journals/cacm/Hoare74}.
Coroutines are not originally established in the object-oriented paradigm.
Object orientation is based on caller-callee interaction, which is asymmetric:
the caller invokes a method in the callee and \emph{blocks} until the
corresponding \emph{return} occurs in the method of the callee.
Object orientation focuses on an object-view with caller-callee interaction and stack-based execution whereas coroutines focus on a process-view with flexible transfer of control.

@@

ABS~\cite{johnsen2012abs,hahnlehjlssw11} is a modeling language for concurrent objects and distributed systems.
ABS uses asynchronous communication among objects.
A message in ABS is bound to a method invocation;
this defines the interface of the sent messages.
In addition, ABS supports future values in its asynchronous communication.
ABS introduces \bfjtt{release} semantics that is based on co-operative scheduling of objects; 
i.e. similar to that of \emph{yield} in coroutines~\cite{creol:broch:owe}.
The ABS semantics is completely formalized by a structural operational semantics~\cite{plotkin:sos,abs:2012}.
This allows ABS models to take advantage of a wide range of static and dynamic analysis techniques~\cite{DBLP:journals/sttt/WongBBGGHMS15,DBLP:conf/cade/DinBH15,DBLP:conf/ifm/GiachinoGLLW13,DBLP:conf/tacas/AlbertAFGGMPR14}.
% 
Co-operative scheduling in ABS has been additionally extended for real-time scheduling with priorities and time constraints~\cite{bjork2013:rtabs,johnsen2012modeling}.
All the above characteristics make the ABS language an attractive choice if it
can be used as a programming language at industrial and business scale.
ABS has mainly developed as a modeling language;
for example, ABS does not support a common I/O API for files or networking,
and it does not provide a standard library of data structures.

JSR 166~\cite{jsr166} introduced a new concurrency model~\cite{DBLP:conf/java/Lea00} in Java 1.5.
The concurrency model lays the grounds to support for the actor model with asynchronous method calls.
This gave rise to various research to support actor model on Java~\cite{actor_frameworks_jvm:agha} and further extend Java as a functional programming language~\cite{odersky1997pizza,henkel2003discovering,nystrom2003polyglot,bracha1998making}.
From an industrial perspective, the development of the Java language from version 6 to version 8 experienced a slow pace that influenced the growth of businesses and justified alternatives.
Scala~\cite{odersky2004scala}, a dynamic, functional, and object-oriented language on JVM, rose to the challenge to fill the gap during the period between Java~7 and Java~8.
Java~8~\cite{jls8} released fundamental features such as lambda expressions that are essential for concurrency and functional programming.

@@

\paragraph{Scalability} 
In distributed systems, the number of messages delivered among objects in the 
environment is not predictable at runtime.
The goal is to ensure the actor system scales in performance with least 
influence from the number of asynchronous messages delivered in the system.
Due to support of co-operative scheduling for asynchronous messages, ABS is a fit for distributed systems.

\paragraph{Separation of Concerns} 
We approach ABS modeling and development with a component-based and modular 
software engineering practices.
The scope of the research spans a number of layers around ABS language:

@@

The core contributions of this thesis target the intersection of object 
orientation, actor model, and concurrency.
We choose Java as the main programming language and as one of the mainstream 
object-oriented languages. 
We formalize a subset of Java and its concurrency API~\cite{jsr166} to 
facilitate formal verification and reasoning about it (cf. Chapter~\ref{ch:p01:ch02}).
We create an abstract mapping from a concurrent-object modeling language, 
ABS~\cite{johnsen2012abs}, to the programming semantics of concurrent Java (cf. Chapter~\ref{ch:p01:ch02}). 
We provide the formal semantics of the mapping and runtime properties of 
the concurrency layer including deadlines and scheduling policies (cf. Chapter~\ref{ch:p01:ch01}).
We provide an implementation of the ABS concurrency layer as a Java API library 
and framework utilizing the latest language additions 
in Java 8~\cite{jsr335:lambda:translation} (cf. Chapter~\ref{ch:p02:ch01}).
We design and implement a runtime monitoring framework, JMSeq, to verify the
the correct ordering of execution of methods through code annotations in JVM (cf. Chapter~\ref{ch:p03:ch02}). 
In addition, we design a large-scale monitoring system as a real-world 
application; the monitoring system is built with ABS concurrent objects 
and formal semantics that leverages schedulability 
analysis to verify correctness of the monitors~\cite{fersman2007task} (cf. Chapter~\ref{ch:p03:ch01}).
Table~\ref{tbl:thesis} summarizes the structure of this text.
